Calibrating the pixel-level Kepler imaging data with a causal data-driven model \\

### 2015-04-02 documents/cpm.tex
typos

\begin{abstract}
Astronomical observations are affected by several kinds of noise, each with its own causal source; 
there is photon noise, stochastic source variability, and residuals coming from imperfect calibration of the detector or telescope. 
The precision of NASA \Kepler\ photometry for exoplanet science ---
the most precise photometric measurements of stars ever made ---
appears to be limited by unknown or untracked variations in spacecraft pointing and temperature, 
and unmodeled stellar variability. Here we present the Causal Pixel Model (\name) for \Kepler\ data, 
a data-driven model intended to capture variability but preserve transit signals. 
The \name\ works at the pixel level; So that
it can capture more fine-grained information about the variation of the spacecraft.
The \name\ predicts each target pixel value from a large number of pixels of other stars sharing the instrument variabilities 
while not containing any information on possible transits in the target star. In addition, we use the target star’s future and past (auto-regression). 
By appropriately separating, for each data point, the data into training and test sets, 
  we ensure that information about any transit will be perfectly isolated from the model. 
The method has four hyper-parameters (the number of predictor stars, the auto-regressive window size, 
and two L2-regularization amplitudes for model components), which we set by cross-validation. 
We determine a generic set of hyper-parameters that works well for most of the stars and apply the method to a corresponding set of target stars. 
We find that we can consistently outperform (for the purposes of exoplanet detection)% 
  the \Kepler\ Pre-search Data Conditioning (PDC) method for exoplanet discovery.

@@

The train-and-test idea comes into the objective function $\chi^2_{m,n}$:
When we are computing the objective function $\chi^2_{m,n}$,
  we are using only time points $n'\in\set{N}_n$ that don't overlap the target (or test) time point $n$.
We train the model---that is, set the parameters $a_{m,n,m'}$ and $b_{m,n,v}$---%
  by choosing the full set of parameters $a_{m,n,m'}$ and $b_{m,n,v}$% 
  that jointly minimize the objective function $\chi^2_{m,n}$.
Fortunately, given the form of the model,
  this minimization is just a linear solve (linear operation on the data).
Importantly however, and perhaps surprisingly, the objective function $\chi^2_{m,n}$ is \emph{different}
  for every target data point (pixel value to be predicted) $I_{m,n}$;
  we have to do an \emph{independent optimization} (or really linear solve)
  of the parameters for every pixel value we want to predict at every time.
That is, every pixel datum $I_{m,n}$ we predict will have
  \emph{different} settings of the parameters $a_{m,n,m'}$ and $b_{m,n,v}$.
Since there are of order $10^{6}$ total pixel values in the \Kepler\ Target Pixel Files for \emph{each \Kepler\ target},
  and of order $10^{3}$ data sources used as basis functions in the linear fitting,
  this represents a lot of linear solves, and of order $10^{9}$ parameters to obtain and record.

@@

Because of the train-and-test framework,  
  we expect the \name\ prediction $S_{n}$ to be a prediction of%
  what the star lightcurve would have under the influence of the spacecraft but with no exoplanet signal 
Based on this idea, 
  we can construct the ``\name\ flux''% 
  (where the inverted commas indicate that it is not a flux in the sense that stellar variability is not preserved)% 
  to be the relative residual between \name\ prediction and SAP Flux $F_{n}$
\begin{eqnarray}
\delta_{n}&\equiv&\frac{F_{n} - S_{n}}{S_{n}}
\quad .
\end{eqnarray} 
In some sense, it is this relative residual--- the \name\ flux that contains the exoplanet transit signals we seek. 
That is, a transit creates a negative residual away from the prediction, 
  and the amount of relative residual is just the fraction of the light eclipsed by the planet (the transit depth).

\subsection{Hyper-parameters}
The \name\ has 4 hyper-parameters--- 
  the number of predictor stars Q (or number of predictor pixels P), 
  the number of autoregressive components S, 
  and the two regularization strength $\lambda_{a}$ and $\lambda_{b}$.
To set these 4 hyper-parameters, 
  in principle we need to run something like a cross-validation on every single star to optimize performance.
But optimizing in this 4-dimensional space is expensive, 
  especially since in the \name\ we need to solve thousands of linear systems, 
  each of which has thousands of free parameters. 
Therefore, in this paper, 
  we just set a general set of ``default'' hyper-parameters ($P=4000$, $Q=3$, $\lambda_a=1e5$, $\lambda_a=1e5$), 
  which works well in most of the stars without high variability. 
To show how this general set of hyper-parameters performs, 
  we present a few comparison between optimized and default hyper-parameters 
  in Fig~\ref{hyperparameter}.\Bernhard{maybe explain somewhere how SNR is calculated}  
We can see that, in Fig~\ref{hyperparameter}, 
  there is less variation in the \name\ flux% 
  when we use optimized hyper-parameters (green points in second panel)% 
  than when we use the default one (red points in the bottom panel). 
In addition, the optimized flux has higher signal to noise ratio of the known transit than the default. 
However, the performance of default hyper-parameters is acceptable 
  and we will use this set of hyper-parameters throughout the rest of the paper 
  to show the general performance of \name.

@@

\section{Discussion}
We have presented a simple yet highly effective method---the \name\ that works in pixel-level 
  based on the causal structure of the \Kepler\ data to calibrate the \Kepler\ lightcurves, 
  which is intended to search exoplanets.
In \name, systematics and stellar variabilities are removed by% 
  either fitting with other stars' lightcures or auto-regressive components,
  while transit signals can still be perfectly preserved by train-and-test framework.
Low variable clean lightcurves can be produced by \name\,  which is ideal for transit planet searching. 
The CDPP noise level estimation is also presented in the paper to show the performance of \name.

@@

One possible way to solve this kind of edge effect 
  but still preserve the transit signals to do the simultaneous fitting 
  for both systematics and transit signal together, 
  like what have been done in \cite{dfm}. 
That is,  
  instead of only modelling the systematics% 
  and then subtracting it from the raw data to produce some de-trended lightcurves,
  a comprehensive model for both all components in the data (systematics, stellar variabilities and transits)  can be made. In this kind of model, there is no need for train-and-test framework,
  since all transit signals have already been modelled 
  and no de-trended lightcurve will be produced. Planets will be found directly in the model itself.

>>>>

\begin{abstract}
Astronomical observations are affected by several kinds of noise, each with its own causal source; 
there is photon noise, stochastic source variability, and residuals coming from imperfect calibration of the detector or telescope. 
The precision of NASA \Kepler\ photometry for exoplanet science ---
the most precise photometric measurements of stars ever made ---
appears to be limited by unknown or untracked variations in spacecraft pointing and temperature, 
and unmodeled stellar variability. Here we present the Causal Pixel Model (\name) for \Kepler\ data, 
a data-driven model intended to capture variability but preserve transit signals. 
The \name\ works at the pixel level; So that
it can capture more fine-grained information about the variation of the spacecraft.
The \name\ predicts each target pixel value from a large number of pixels of other stars sharing the instrument variabilities 
while not containing any information on possible transits in the target star. In addition, we use the target star’s future and past (auto-regression). 
By appropriately separating, for each data point, the data into training and test sets, 
  we ensure that information about any transit will be perfectly isolated from the model. 
The method has four hyper-parameters (the number of predictor stars, the auto-regressive window size, 
and two L2-regularization amplitudes for model components), which we set by cross-validation. 
We determine a generic set of hyper-parameters that works well for most of the stars and apply the method to a corresponding set of target stars. 
We find that we can consistently outperform (for the purposes of exoplanet detection) 
  the \Kepler\ Pre-search Data Conditioning (PDC) method for exoplanet discovery.

@@

The train-and-test idea comes into the objective function $\chi^2_{m,n}$:
When we are computing the objective function $\chi^2_{m,n}$,
  we are using only time points $n'\in\set{N}_n$ that don't overlap the target (or test) time point $n$.
We train the model---that is, set the parameters $a_{m,n,m'}$ and $b_{m,n,v}$---%
  by choosing the full set of parameters $a_{m,n,m'}$ and $b_{m,n,v}$ 
  that jointly minimize the objective function $\chi^2_{m,n}$.
Fortunately, given the form of the model,
  this minimization is just a linear solve (linear operation on the data).
Importantly however, and perhaps surprisingly, the objective function $\chi^2_{m,n}$ is \emph{different}
  for every target data point (pixel value to be predicted) $I_{m,n}$;
  we have to do an \emph{independent optimization} (or really linear solve)
  of the parameters for every pixel value we want to predict at every time.
That is, every pixel datum $I_{m,n}$ we predict will have
  \emph{different} settings of the parameters $a_{m,n,m'}$ and $b_{m,n,v}$.
Since there are of order $10^{6}$ total pixel values in the \Kepler\ Target Pixel Files for \emph{each \Kepler\ target},
  and of order $10^{3}$ data sources used as basis functions in the linear fitting,
  this represents a lot of linear solves, and of order $10^{9}$ parameters to obtain and record.

@@

Because of the train-and-test framework,  
  we expect the \name\ prediction $S_{n}$ to be a prediction of
  what the star lightcurve would have under the influence of the spacecraft but with no exoplanet signal 
Based on this idea, 
  we can construct the ``\name\ flux'' 
  (where the inverted commas indicate that it is not a flux in the sense that stellar variability is not preserved) 
  to be the relative residual between \name\ prediction and SAP Flux $F_{n}$
\begin{eqnarray}
\delta_{n}&\equiv&\frac{F_{n} - S_{n}}{S_{n}}
\quad .
\end{eqnarray} 
In some sense, it is this relative residual--- the \name\ flux that contains the exoplanet transit signals we seek. 
That is, a transit creates a negative residual away from the prediction, 
  and the amount of relative residual is just the fraction of the light eclipsed by the planet (the transit depth).

\subsection{Hyper-parameters}
The \name\ has 4 hyper-parameters--- 
  the number of predictor stars Q (or number of predictor pixels P), 
  the number of autoregressive components S, 
  and the two regularization strength $\lambda_{a}$ and $\lambda_{b}$.
To set these 4 hyper-parameters, 
  in principle we need to run something like a cross-validation on every single star to optimize performance.
But optimizing in this 4-dimensional space is expensive, 
  especially since in the \name\ we need to solve thousands of linear systems, 
  each of which has thousands of free parameters. 
Therefore, in this paper, 
  we just set a general set of ``default'' hyper-parameters ($P=4000$, $Q=3$, $\lambda_a=1e5$, $\lambda_a=1e5$), 
  which works well in most of the stars without high variability. 
To show how this general set of hyper-parameters performs, 
  we present a few comparison between optimized and default hyper-parameters 
  in Fig~\ref{hyperparameter}.\Bernhard{maybe explain somewhere how SNR is calculated}  
We can see that, in Fig~\ref{hyperparameter}, 
  there is less variation in the \name\ flux 
  when we use optimized hyper-parameters (green points in second panel) 
  than when we use the default one (red points in the bottom panel). 
In addition, the optimized flux has higher signal to noise ratio of the known transit than the default. 
However, the performance of default hyper-parameters is acceptable 
  and we will use this set of hyper-parameters throughout the rest of the paper 
  to show the general performance of \name.

@@

\section{Discussion}
We have presented a simple yet highly effective method---the \name\ that works in pixel-level 
  based on the causal structure of the \Kepler\ data to calibrate the \Kepler\ lightcurves, 
  which is intended to search exoplanets.
In \name, systematics and stellar variabilities are removed by
  either fitting with other stars' lightcures or auto-regressive components,
  while transit signals can still be perfectly preserved by train-and-test framework.
Low variable clean lightcurves can be produced by \name\,  which is ideal for transit planet searching. 
The CDPP noise level estimation is also presented in the paper to show the performance of \name.

@@

One possible way to solve this kind of edge effect 
  but still preserve the transit signals is to do the simultaneous fitting 
  for both systematics and transit signal together, 
  like what have been done in \cite{dfm}. 
That is,  
  instead of only modelling the systematics 
  and then subtracting it from the raw data to produce some de-trended lightcurves,
  a comprehensive model for all components in the data (systematics, stellar variabilities and transits)  can be made. In this kind of model, there is no need for train-and-test framework,
  since all transit signals have already been modelled 
  and no de-trended lightcurve will be produced. Planets will be found directly in the model itself.
